---
title: "Naive Bayes Classification"
output: html_document
---
#Loading Data

First load Libraries and assign paths to spam or ham folders in WD. Spam and Ham messages have been downloaded from 
Apache's Spam Assassian public corpus. 
```{r, results="hide"}
library(tm)
library(stringr)
library(plyr)
library(kernlab)

spam.path <-"spam/"
spam2.path <- "spam_2/"
easyham.path <- "easy_ham/"
easyham2.path <- "easy_ham_2/"
hardham.path <- "hard_ham/"
hardham2.path <- "hard_ham 2/"
```

# Support Functions

This Function returns a single element vector of just the email body. Here we are only using words as features. Importantly, the body of the message always begins after the first full line break (the rest is just meta data).
```{r}
get.msg <- function(path)
{
  con <- file(path, open = "rt", encoding = "latin1")
  text <- readLines(con)
  msg <- text[seq(which(text == "")[1] + 1, length(text), 1)]
  close(con)
  return(paste(msg, collapse = "\n"))
}
```


This function creates a Term Document Matrix (TDM) from the corpus of SPAM email. We remove all stopwords, punctions, numbers, and only include documents that occur more than twice. This TDM is used to create the feature set used to do train our classifier.
```{r}
get.tdm <- function(doc.vec)
{
  doc.corpus <- Corpus(VectorSource(doc.vec))
  control<-list(stopwords = TRUE, removePunctuation = TRUE,removeNumbers = TRUE, minDocFreq = 2)
  doc.dtm <- TermDocumentMatrix(doc.corpus, control)
  return(doc.dtm)
}
```


This function takes a file path to an email file and a string( the term parameter), and returns the count of that term in the email body.
```{r}
count.word <- function(path, term)
{
  msg <- get.msg(path)
  msg.corpus <- Corpus(VectorSource(msg))
  control<-list(stopwords = TRUE, removePunctuation = TRUE,removeNumbers = TRUE, minDocFreq = 2)
  msg.tdm <- TermDocumentMatrix(msg.corpus, control)
  word.freq <- rowSums(as.matrix(msg.tdm))
  term.freq <- word.freq[which(names(word.freq) == term)]
  return(ifelse(length(term.freq) > 0, term.freq, 0))
}
```


This is our primary classification function for spam vs ham. It takes a parameter for the path to the file to be classified and a dataframe of the trained data. It also takes a prior distribution paramater (see report), and a value c which is the probablity words that are not in the training data appear in the email. The functions returns the naive bayes probability that the email is spam. The prior here can be adjusted (default is 50% or an even distribution of spam to ham), but after research we have determined the most accurate distribution of incoming emails is 80% Ham to 20% Spam. 
```{r}
classify.email <- function(path, training.df, prior = 0.5, c = 1e-6)
{
  msg <- get.msg(path)
  msg.tdm <- get.tdm(msg)
  msg.freq <- rowSums(as.matrix(msg.tdm))
  # Find intersections of words
  msg.match <- intersect(names(msg.freq), training.df$term)
  # Perform the naive Bayes calculation
  if(length(msg.match) < 1)
  {
    return(prior * c ^ (length(msg.freq)))
  }
  else
  {
    match.probs <- training.df$occurrence[match(msg.match, training.df$term)]
    return(prior * prod(match.probs) * c ^ (length(msg.freq) - length(msg.match)))
  }
}
```


```{r}
get.results <- function(bool.vector)
{
  results <- c(length(bool.vector[which(bool.vector == FALSE)]) / length(bool.vector),
               length(bool.vector[which(bool.vector == TRUE)]) / length(bool.vector))
  return(results)
}
```

#Create training data

To create training data, all spam emails are read into a single vector then a Document Term matrix is generate to provide feature set for training data. From the Document Term Matrix the term density and occurance rates are calculated. This is done indentically to create training sets from spam and ham, but only the code for spam is shown below.
```{r,message='false'}
# Get all the SPAM email into a single vector
spam.docs <- dir(spam.path)
spam.docs <- spam.docs[which(spam.docs != "cmds")]
all.spam <- sapply(spam.docs,
                   function(p) get.msg(file.path(spam.path, p)))

# Create a DocumentTermMatrix from that vector
spam.tdm <- get.tdm(all.spam)

# Create a data frame that provides the feature set from the training SPAM data
spam.matrix <- as.matrix(spam.tdm)
spam.counts <- rowSums(spam.matrix)
spam.df <- data.frame(cbind(names(spam.counts),
                            as.numeric(spam.counts)),
                      stringsAsFactors = FALSE)
names(spam.df) <- c("term", "frequency")
spam.df$frequency <- as.numeric(spam.df$frequency)
spam.occurrence <- sapply(1:nrow(spam.matrix),
                          function(i)
                          {
                            length(which(spam.matrix[i, ] > 0)) / ncol(spam.matrix)
                          })
spam.density <- spam.df$frequency / sum(spam.df$frequency)

# Add the term density and occurrence rate
spam.df <- transform(spam.df,
                     density = spam.density,
                     occurrence = spam.occurrence)
```



```{r,include='false',message='false'}
#Do same for Ham
easyham.docs <- dir(easyham.path)
easyham.docs <- easyham.docs[which(easyham.docs != "cmds")]
all.easyham <- sapply(easyham.docs[1:length(spam.docs)],
                      function(p) get.msg(file.path(easyham.path, p)))


easyham.tdm <- get.tdm(all.easyham)

easyham.matrix <- as.matrix(easyham.tdm)
easyham.counts <- rowSums(easyham.matrix)
easyham.df <- data.frame(cbind(names(easyham.counts),
                               as.numeric(easyham.counts)),
                         stringsAsFactors = FALSE)
names(easyham.df) <- c("term", "frequency")
easyham.df$frequency <- as.numeric(easyham.df$frequency)
easyham.occurrence <- sapply(1:nrow(easyham.matrix),
                             function(i)
                             {
                               length(which(easyham.matrix[i, ] > 0)) / ncol(easyham.matrix)
                             })
easyham.density <- easyham.df$frequency / sum(easyham.df$frequency)

easyham.df <- transform(easyham.df,
                        density = easyham.density,
                        occurrence = easyham.occurrence)




# Run classifer against HARD HAM
hardham.docs <- dir(hardham.path)
hardham.docs <- hardham.docs[which(hardham.docs != "cmds")]

hardham.spamtest <- sapply(hardham.docs,
                           function(p) classify.email(file.path(hardham.path, p), training.df = spam.df))

hardham.hamtest <- sapply(hardham.docs,
                          function(p) classify.email(file.path(hardham.path, p), training.df = easyham.df))

hardham.res <- ifelse(hardham.spamtest > hardham.hamtest,
                      TRUE,
                      FALSE)
summary(hardham.res)
```

#Classification 

Here I adjust prior the classifcation function to better reflect the true distribution of spam vs ham. Research has indicated that the distribution is approximentaly 80% Ham and 20% spam. To make change the prior disribution assumption for this naive bayes classifier, adjust the prior parameter on the spam.classifier function
```{r}
spam.classifier <- function(path)
{
  pr.spam <- classify.email(path, spam.df,prior=.2)
  pr.ham <- classify.email(path, easyham.df,prior=.8)
  #returns spam if prob(spam) > prob(ham)
  return(c(pr.spam, pr.ham, ifelse(pr.spam > pr.ham, 1, 0)))
}
```

#### Impliment support functions to conduct classification task
Get List of spam, and all ham emails. Delete email files ending in cmds because they don't contain email text data.
```{r}
easyham2.docs <- dir(easyham2.path)
easyham2.docs <- easyham2.docs[which(easyham2.docs != "cmds")]

hardham2.docs <- dir(hardham2.path)
hardham2.docs <- hardham2.docs[which(hardham2.docs != "cmds")]

spam2.docs <- dir(spam2.path)
spam2.docs <- spam2.docs[which(spam2.docs != "cmds")]
```

The code block impliments the support functions to classify the easy ham,hard ham, and spam emails. The lapply function applied the spam.classifier support function to every email in the directory. The anonymous function passed through lapply allows us to concentate the emails by rows as it cycles through each email, the value p passed through the anonymous function acts as a place holder in the file.path list. 

```{r}
easyham2.class <- suppressWarnings(lapply(easyham2.docs,
                                          function(p)
                                          {
                                            spam.classifier(file.path(easyham2.path, p))
                                          }))
hardham2.class <- suppressWarnings(lapply(hardham2.docs,
                                          function(p)
                                          {
                                            spam.classifier(file.path(hardham2.path, p))
                                          }))
spam2.class <- suppressWarnings(lapply(spam2.docs,
                                       function(p)
                                       {
                                         spam.classifier(file.path(spam2.path, p))
                                       }))

```


The next step is to create a single dataframe by merging the spam, hard and easy ham classifcations into one data frame. The do.call functions calls rbind to hold the argumnts of the rbind functions, and in combination with the cbind function it effectively creates a final classification matrix. 
```{r}
easyham2.matrix <- do.call(rbind, easyham2.class)
easyham2.final <- cbind(easyham2.matrix, "EASYHAM")

hardham2.matrix <- do.call(rbind, hardham2.class)
hardham2.final <- cbind(hardham2.matrix, "HARDHAM")

spam2.matrix <- do.call(rbind, spam2.class)
spam2.final <- cbind(spam2.matrix, "SPAM")

class.matrix <- rbind(easyham2.final, hardham2.final, spam2.final)
class.df <- data.frame(class.matrix, stringsAsFactors = FALSE)
names(class.df) <- c("Pr.SPAM" ,"Pr.HAM", "Class", "Type")
class.df$Pr.SPAM <- as.numeric(class.df$Pr.SPAM)
class.df$Pr.HAM <- as.numeric(class.df$Pr.HAM)
class.df$Class <- as.logical(as.numeric(class.df$Class))
class.df$Type <- as.factor(class.df$Type)
```



```{r,echo="false"}

easyham2.col <- get.results(subset(class.df, Type == "EASYHAM")$Class)
hardham2.col <- get.results(subset(class.df, Type == "HARDHAM")$Class)
spam2.col <- get.results(subset(class.df, Type == "SPAM")$Class)

class.res <- rbind(easyham2.col, hardham2.col, spam2.col)
colnames(class.res) <- c("NOT SPAM", "SPAM")
class.res[c(1,2),]
```


#### ROC Curve

Plot an ROC curve for evaluation
```{r}
library(ROCR)


#Get pr(spam) for everything

ROC.Testing<- class.df[!c(class.df$Type == "EASYHAM"),]
ROC.Testing$Pr.SPAM





ROC.Testing$Type1= as.integer(ROC.Testing$Type)
ROC.Testing$Type1= ROC.Testing$Type1-2
ROC.Testing$Type1 #1 is spam


pred = prediction(ROC.Testing$Pr.SPAM,ROC.Testing$Type1)
perf= performance(pred,measure="tpr",x.measure="fpr")
plot(perf,col=rainbow(10),colorize=T)
```


